---
title: "Indep_project"
author: "Aaron Blue and James Kosmopoulos"
output: html_document
---

## load required packages

```{r packages}
library(tidyverse) ## reshaping spreadsheets
library(lme4) ## linear mixed models
library(car) ## Anova function
```


## load data

```{r load_data}
metadata <- read.csv("./metadata.csv")
```


## summarize data


```{r}
metadata.long <- gather(metadata, key = "trait", value = "value", nodulenum, meannodulemass, chloro, shootmass)

plants.year.sum <- metadata.long %>% group_by(trait, plantspori, dataset_year, nitrogen) %>% summarize(count = n(),
            mean = mean(value, na.rm = TRUE), 
            SE = sd(value, na.rm = TRUE)/sqrt(count))

strains.year.sum <- metadata.long %>% group_by(trait, strain, dataset_year, nitrogen) %>% summarize(count = n(),
            mean = mean(value, na.rm = TRUE), 
            SE = sd(value, na.rm = TRUE)/sqrt(count))

plants.nitrogen.sum <- metadata.long %>% group_by(trait, plantspori, nitrogen) %>% summarize(count = n(),
            mean = mean(value, na.rm = TRUE), 
            SE = sd(value, na.rm = TRUE)/sqrt(count))

strains.nitrogen.sum <- metadata.long %>% group_by(trait, strain, nitrogen) %>% summarize(count = n(),
            mean = mean(value, na.rm = TRUE), 
            SE = sd(value, na.rm = TRUE)/sqrt(count))
```


## Mixed effects models

```{r}
lmm_shoot_2008 <- lmer(shootmass ~ plantspori + nitrogen + plantspori:nitrogen + (1|strain) + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2008"))
summary(lmm_shoot_2008)
Anova(lmm_shoot_2008, type = 3) # nitrogen has sig effect, make plot
plot(lmm_shoot_2008)

lmm_shoot_2018 <- lmer(shootmass ~ plantspori + nitrogen + plantspori:nitrogen + (1|strain) + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2018"))
summary(lmm_shoot_2018)
Anova(lmm_shoot_2018, type = 3) # nothing is sig
plot(lmm_shoot_2018)

lmm_chloro_2008 <- lmer(log(chloro) ~ plantspori + nitrogen + plantspori:nitrogen + (1|strain) + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2008"))
summary(lmm_chloro_2008)
Anova(lmm_chloro_2008, type = 3) # nitrogen has sig effect, make plot
plot(lmm_chloro_2008)

lmm_chloro_2018 <- lmer(log(chloro) ~ plantspori + nitrogen + plantspori:nitrogen + (1|strain) + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2018"))
summary(lmm_chloro_2018)
Anova(lmm_chloro_2018, type = 3) # plant species and nitrogen has sig effect, make plot
plot(lmm_chloro_2018)

lmm_nodulenum_2008 <- lmer(sqrt(nodulenum) ~ plantspori + nitrogen + plantspori:nitrogen + (1|strain) + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2008"))
summary(lmm_nodulenum_2008)
Anova(lmm_nodulenum_2008, type = 3) # nothing is sig
plot(lmm_nodulenum_2008)

lmm_nodulenum_2018 <- lmer(sqrt(nodulenum) ~ plantspori + nitrogen + plantspori:nitrogen + (1|strain) + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2018"))
summary(lmm_nodulenum_2018)
Anova(lmm_nodulenum_2018, type = 3) # nothing is sig
plot(lmm_nodulenum_2018)

lmm_meannodulemass_2008 <- lmer(sqrt(meannodulemass) ~ plantspori + nitrogen + plantspori:nitrogen + (1|strain) + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2008"))
summary(lmm_meannodulemass_2008)
Anova(lmm_meannodulemass_2008, type = 3)  # nothing is sig
plot(lmm_meannodulemass_2008)

lmm_meannodulemass_2018 <- lmer(sqrt(meannodulemass) ~ plantspori + nitrogen + plantspori:nitrogen + (1|strain) + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2018"))
summary(lmm_meannodulemass_2018)
Anova(lmm_meannodulemass_2018, type = 3)  # nothing is sig
plot(lmm_meannodulemass_2018)
```

#Heritability Function Prep

```{r}
chloro_2008 <- lmer(log(chloro) ~ plantspori + nitrogen + plantspori:nitrogen + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2008"))

nodulenum_2008 <- lmer(sqrt(nodulenum) ~ plantspori + nitrogen + plantspori:nitrogen + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2008"))

shootmass_2008 <- lmer(shootmass ~ plantspori + nitrogen + plantspori:nitrogen + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2008"))

meannodulemass_2008 <- lmer(sqrt(meannodulemass) ~ plantspori + nitrogen + plantspori:nitrogen + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2008"))

chloro_2018 <- lmer(log(chloro) ~ plantspori + nitrogen + plantspori:nitrogen + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2018"))

nodulenum_2018 <- lmer(sqrt(nodulenum) ~ plantspori + nitrogen + plantspori:nitrogen + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2018"))

shootmass_2018 <- lmer(shootmass ~ plantspori + nitrogen + plantspori:nitrogen + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2018"))

meannodulemass_2018 <- lmer(sqrt(meannodulemass) ~ plantspori + nitrogen + plantspori:nitrogen + (1|harvestdate) + (1|harvester), data = metadata %>% filter(dataset_year=="2018"))


```

#Heritability Function Building
```{r}


Heritability_Function <- function(model, ResponseVariable_Year){
   
lmm_x_2008  <- model
lmm_x_2008_dplot <- ResponseVariable_Year

# model comparison
comp_chloro_2008 <- anova(lmm_x_2008, lmm_x_2008_dplot)
var_chloro_2008 <- VarCorr(lmm_x_2008)

# test results
chi_chloro_2008 <- comp_chloro_2008$Chisq
Df_chloro_2008 <- comp_chloro_2008$Df
sig_chloro_2008<- (comp_chloro_2008$`Pr(>Chisq)`)/2

# heritability
Var_chloro_2008 <- as.data.frame(var_chloro_2008)
Var_chloro_2008 <- Var_chloro_2008 %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_chloro_2008 <- Var_chloro_2008[1,7]


# combine stats and heritability
comb_chloro_2008 <- cbind(chi_chloro_2008[2], Df_chloro_2008[2], sig_chloro_2008[2], VG_chloro_2008)
colnames(comb_chloro_2008) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")   
   
model <- comb_chloro_2008

return(model)
}


```

#Completed Heritability Function in this format: (Model, ResponseVariable_Year)

```{r}

Heritability_Function(lmm_chloro_2008, chloro_2008)

```


## Plot results from models

```{r}
## By plant species
# Shoot biomass 2018, nitrogen significant, slight GxE
(plot.shootmass.2018 <- ggplot(data = plants.year.sum %>% filter(trait=="shootmass", dataset_year=="2018"), aes(x=nitrogen, y=mean, group=plantspori, color=plantspori)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Environment") +
   ylab("Shoot biomass") +
   theme_bw()
  )

# Shoot biomass 2008, not significant
(plot.shootmass.2008 <- ggplot(data = plants.year.sum %>% filter(trait=="shootmass", dataset_year=="2008"), aes(x=nitrogen, y=mean, group=plantspori, color=plantspori)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Environment") +
   ylab("Shoot biomass") +
   theme_bw()
  )

# Chloro 2018, nitrgen and plant species significant, slight GxE
(plot.chloro.2018 <- ggplot(data = plants.year.sum %>% filter(trait=="chloro", dataset_year=="2018"), aes(x=nitrogen, y=mean, group=plantspori, color=plantspori)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Environment") +
   ylab("Chlorophyll content") +
   theme_bw()
  )

# Chloro 2008, nitrogen significant, weak GxE?
(plot.chloro.2008 <- ggplot(data = plants.year.sum %>% filter(trait=="chloro", dataset_year=="2008"), aes(x=nitrogen, y=mean, group=plantspori, color=plantspori)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Environment") +
   ylab("Chlorophyll content") +
   theme_bw()
  )

```

## We cannot plot reaction norms grouped by strain because strains originally isolated from low nitrogen plots (control) are not identical to strains isolated form fertilized plots (nitrogen). We could do this before because we were using "fert" which includes both strains, but we can't use fert anymore and we had to use nitrogen. I think this is okay though since we don't include strain as a fixed effect, we might just have to explain why though.

## Calculate heritability

```{r}
# We need to make a function to calculate heritability from out models or do it from scratch like before
```





## reaction norms


# nodulenum

```{r}
## by year
(plot.nodulenum.year<- ggplot(data = strains.year.sum %>% filter(trait=="nodulenum"), aes(x=dataset_year, y=mean, group=strain, color=strain)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Year") +
   ylab("Nodule number") +
   theme_bw() +
   theme(legend.position = "none")
  )

# by environment (James help how do I do this the strains arent the same)

(plot.nodulenum.env <- ggplot(data = strains.nitrogen.sum %>% filter(trait=="nodulenum"), aes(x=nitrogen, y=mean, group=strain, color=strain)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Year") +
   ylab("Nodule number") +
   theme_bw() +
   theme(legend.position = "none")
  )


```

# Nodmass

```{r}
## by year
(plot.nodmass.year <- ggplot(data = strains.year.sum %>% filter(trait=="meannodulemass"), aes(x=dataset_year, y=mean, group=strain, color=strain)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Year") +
   ylab("Nodule mass") +
   theme_bw() +
   theme(legend.position = "none")
  )

## by environment
(plot.nodmass.env <- ggplot(data = strains.nitrogen.sum %>% filter(trait=="meannodulemass"), aes(x=nitrogen, y=mean, group=strain, color=strain)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Environment") +
   ylab("Nodule mass") +
   theme_bw() +
   theme(legend.position = "none")
  )

```

# Chloro

```{r}
## by year
(plot.chloro.year <- ggplot(data = plants.year.sum %>% filter(trait=="chloro"), aes(x=dataset_year, y=mean, group=plantspori, color=plantspori)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Year") +
   ylab("Chlorophyll content") +
   theme_bw()
  )

## by N environment
(plot.chloro.env <- ggplot(data = plants.nitrogen.sum %>% filter(trait=="chloro"), aes(x=nitrogen, y=mean, group=plantspori, color=plantspori)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Environment") +
   ylab("Chlorophyll content") +
   theme_bw()
  )

```

# Shoot mass

```{r}
## by year
(plot.shootmass.year <- ggplot(data = plants.year.sum %>% filter(trait=="shootmass"), aes(x=dataset_year, y=mean, group=plantspori, color=plantspori)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Year") +
   ylab("Shoot biomass") +
   theme_bw()
  )

## by N environment
(plot.shootmass.env <- ggplot(data = plants.nitrogen.sum %>% filter(trait=="shootmass"), aes(x=nitrogen, y=mean, group=plantspori, color=plantspori)) +
   geom_point(size=3, position = position_dodge(0.3)) +
   geom_line(position = position_dodge(0.3)) +
   xlab("Environment") +
   ylab("Shoot biomass") +
   theme_bw()
  )
```


## Plant shoot biomass models

```{r}
# test if shoot biomass is normally distributed, p<0.05 means data is not normally distributed and we must transform our data with sqrt or log
shapiro.test(metadata$shootmass) # p<0.05, transform data

```


# Shoot mass heritability by year

```{r}
# 2008
lmm_shoot_2008 <- lmer(sqrt(shootmass) ~ dataset_year + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2008"))
Anova(lmm_shoot_2008, type = 3)

# full model minus strain random effect
lmm_shoot_2008_dplot <- lmer(sqrt(shootmass) ~ dataset_year + (1|harvester) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2008"))


# model comparison
comp_shoot_2008 <- anova(lmm_shoot_2008, lmm_shoot_2008_dplot)
var_shoot_2008 <- VarCorr(lmm_shoot_2008)

# test results
chi_shoot_2008 <- comp_shoot_2008$Chisq
Df_shoot_2008 <- comp_shoot_2008$Df
sig_shoot_2008<- (comp_shoot_2008$`Pr(>Chisq)`)/2

# heritability
Var_shoot_2008 <- as.data.frame(var_shoot_2008)
Var_shoot_2008 <- Var_shoot_2008 %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_shoot_2008 <- Var_shoot_2008[1,7]


# combine stats and heritability
comb_shoot_2008 <- cbind(chi_shoot_2008[2], Df_shoot_2008[2], sig_shoot_2008[2], VG_shoot_2008)
colnames(comb_shoot_2008) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")
```

```{r}
# 2018
lmm_shoot_2018 <- lmer(sqrt(shootmass) ~  dataset_year + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2018"))
Anova(lmm_shoot_2018, type = 3)

# full model minus strain random effect
lmm_shoot_2018_dplot <- lmer(sqrt(shootmass) ~ dataset_year + (1|harvester) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2018"))


# model comparison
comp_shoot_2018 <- anova(lmm_shoot_2018, lmm_shoot_2018_dplot)
var_shoot_2018 <- VarCorr(lmm_shoot_2018)

# test results
chi_shoot_2018 <- comp_shoot_2018$Chisq
Df_shoot_2018 <- comp_shoot_2018$Df
sig_shoot_2018<- (comp_shoot_2018$`Pr(>Chisq)`)/2

# heritability
Var_shoot_2018 <- as.data.frame(var_shoot_2018)
Var_shoot_2018 <- Var_shoot_2018 %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_shoot_2018 <- Var_shoot_2018[1,7]


# combine stats and heritability
comb_shoot_2018 <- cbind(chi_shoot_2018[2], Df_shoot_2018[2], sig_shoot_2018[2], VG_shoot_2018)
colnames(comb_shoot_2018) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")

```

# Shoot mass heritability by N environment

```{r}
# Control
# full model with all random effects
lmm_mass_0N <- lmer(sqrt(shootmass) ~  dataset_year  + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(nitrogen=="Control"))
Anova(lmm_mass_0N, type = 3)
summary(lmm_mass_0N)

# model without plot as a random effect
lmm_mass_0N_dplot <- lmer(sqrt(shootmass) ~  dataset_year  + (1|harvester) + (1|harvestdate), data = metadata %>% filter(nitrogen =="Control"))

# model comparison
comp_mass_0N <- anova(lmm_mass_0N, lmm_mass_0N_dplot)
var_mass_0N <- VarCorr(lmm_mass_0N)

# test results
chi_mass_0N <- comp_mass_0N$Chisq
Df_mass_0N <- comp_mass_0N$Df
sig_mass_0N <- (comp_mass_0N$`Pr(>Chisq)`)/2

# heritability
Var_mass_0N <- as.data.frame(var_mass_0N)
Var_mass_0N <- Var_mass_0N %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_mass_0N <- Var_mass_0N[1,7]


# combine stats and heritability
comb_mass_0N <- cbind(chi_mass_0N[2], Df_mass_0N[2], sig_mass_0N[2], VG_mass_0N)
colnames(comb_mass_0N) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")

```

```{r}


# N-fertilized
# full model with all random effects
lmm_mass_fertilized <- lmer(sqrt(shootmass) ~  dataset_year  + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(nitrogen=="Nitrogen"))
Anova(lmm_mass_fertilized, type = 3)
summary(lmm_mass_fertilized)

# model without plot as a random effect
lmm_mass_fertilized_dplot <- lmer(sqrt(shootmass) ~  dataset_year  + (1|harvester) + (1|harvestdate), data = metadata %>% filter(nitrogen =="Nitrogen"))

# model comparison
comp_mass_fertilized <- anova(lmm_mass_fertilized, lmm_mass_fertilized_dplot)
var_mass_fertilized <- VarCorr(lmm_mass_fertilized)

# test results
chi_mass_fertilized <- comp_mass_fertilized$Chisq
Df_mass_fertilized <- comp_mass_fertilized$Df
sig_mass_fertilized <- (comp_mass_fertilized$`Pr(>Chisq)`)/2

# heritability
Var_mass_fertilized <- as.data.frame(var_mass_fertilized)
Var_mass_fertilized <- Var_mass_fertilized %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_mass_fertilized <- Var_mass_fertilized[1,7]


# combine stats and heritability
comb_mass_fertilized <- cbind(chi_mass_fertilized[2], Df_mass_fertilized[2], sig_mass_fertilized[2], VG_mass_fertilized)
colnames(comb_mass_fertilized) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")
```





## Plant chlorophyll models

```{r}
# test if shoot biomass is normally distributed
shapiro.test(metadata$chloro) # p<0.05, transform data

```

# Chloro heritability by year

```{r}
# 2008
lmm_chloro_2008 <- lmer(sqrt(chloro) ~ nitrogen + leafnum + nitrogen:leafnum +(1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2008"))
Anova(lmm_chloro_2008, type = 3)

# full model minus strain random effect
lmm_chloro_2008_dplot <- lmer(sqrt(chloro) ~ nitrogen + leafnum + nitrogen:leafnum + (1|harvester) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2008"))


# model comparison
comp_chloro_2008 <- anova(lmm_chloro_2008, lmm_chloro_2008_dplot)
var_chloro_2008 <- VarCorr(lmm_chloro_2008)

# test results
chi_chloro_2008 <- comp_chloro_2008$Chisq
Df_chloro_2008 <- comp_chloro_2008$Df
sig_chloro_2008<- (comp_chloro_2008$`Pr(>Chisq)`)/2

# heritability
Var_chloro_2008 <- as.data.frame(var_chloro_2008)
Var_chloro_2008 <- Var_chloro_2008 %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_chloro_2008 <- Var_chloro_2008[1,7]


# combine stats and heritability
comb_chloro_2008 <- cbind(chi_chloro_2008[2], Df_chloro_2008[2], sig_chloro_2008[2], VG_chloro_2008)
colnames(comb_chloro_2008) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")

```

```{r}
#2018
lmm_chloro_2018 <- lmer(sqrt(chloro) ~  leafnum + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2018"))
Anova(lmm_chloro_2018, type = 3)

# full model minus strain random effect
lmm_chloro_2018_dplot <- lmer(sqrt(chloro) ~  leafnum + (1|harvester) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2018"))


# model comparison
comp_chloro_2018 <- anova(lmm_chloro_2018, lmm_chloro_2018_dplot)
var_chloro_2018 <- VarCorr(lmm_chloro_2018)

# test results
chi_chloro_2018 <- comp_chloro_2018$Chisq
Df_chloro_2018 <- comp_chloro_2018$Df
sig_chloro_2018<- (comp_chloro_2018$`Pr(>Chisq)`)/2

# heritability
Var_chloro_2018 <- as.data.frame(var_chloro_2018)
Var_chloro_2018 <- Var_chloro_2018 %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_chloro_2018 <- Var_chloro_2018[1,7]


# combine stats and heritability
comb_chloro_2018 <- cbind(chi_chloro_2018[2], Df_chloro_2018[2], sig_chloro_2018[2], VG_chloro_2018)
colnames(comb_chloro_2018) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")
```


# Chloro heritability by N environment

```{r}
# 0N
lmm_chloro_0N <- lmer(sqrt(chloro) ~  dataset_year  + leafnum + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(nitrogen=="Control"))
Anova(lmm_chloro_0N, type = 3)
summary(lmm_chloro_0N)

# model without plot as a random effect
lmm_chloro_0N_dplot <- lmer(sqrt(chloro) ~  dataset_year  + leafnum + (1|harvester) + (1|harvestdate), data = metadata %>% filter(nitrogen =="Control"))

# model comparison
comp_chloro_0N <- anova(lmm_chloro_0N, lmm_chloro_0N_dplot)
var_chloro_0N <- VarCorr(lmm_chloro_0N)

# test results
chi_chloro_0N <- comp_chloro_0N$Chisq
Df_chloro_0N <- comp_chloro_0N$Df
sig_chloro_0N <- (comp_chloro_0N$`Pr(>Chisq)`)/2

# heritability
Var_chloro_0N <- as.data.frame(var_chloro_0N)
Var_chloro_0N <- Var_chloro_0N %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_chloro_0N <- Var_chloro_0N[1,7]


# combine stats and heritability
comb_chloro_0N <- cbind(chi_chloro_0N[2], Df_chloro_0N[2], sig_chloro_0N[2], VG_chloro_0N)
colnames(comb_chloro_0N) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")

```

```{r}


# N-fertilized
# full model with all random effects
lmm_chloro_fertilized <- lmer(sqrt(chloro) ~  dataset_year  + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(nitrogen=="Nitrogen"))
Anova(lmm_chloro_fertilized, type = 3)
summary(lmm_chloro_fertilized)

# model without plot as a random effect
lmm_chloro_fertilized_dplot <- lmer(sqrt(chloro) ~  dataset_year  + (1|harvester) + (1|harvestdate), data = metadata %>% filter(nitrogen =="Nitrogen"))

# model comparison
comp_chloro_fertilized <- anova(lmm_chloro_fertilized, lmm_chloro_fertilized_dplot)
var_chloro_fertilized <- VarCorr(lmm_chloro_fertilized)

# test results
chi_chloro_fertilized <- comp_chloro_fertilized$Chisq
Df_chloro_fertilized <- comp_chloro_fertilized$Df
sig_chloro_fertilized <- (comp_chloro_fertilized$`Pr(>Chisq)`)/2

# heritability
Var_chloro_fertilized <- as.data.frame(var_chloro_fertilized)
Var_chloro_fertilized <- Var_chloro_fertilized %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_chloro_fertilized <- Var_chloro_fertilized[1,7]


# combine stats and heritability
comb_chloro_fertilized <- cbind(chi_chloro_fertilized[2], Df_chloro_fertilized[2], sig_chloro_fertilized[2], VG_chloro_fertilized)
colnames(comb_chloro_fertilized) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")
```



## Rhizobium nodulenum models

```{r}
# test if shoot biomass is normally distributed
shapiro.test(metadata$nodulenum) # p<0.05, transform data
```


# nodulenum heritability by year

```{r}
# 2008
# full model with all random effects
lmm_nod_2008 <- lmer(sqrt(nodulenum) ~ plantspori + plant_age + plantspori:plant_age + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2008"))
Anova(lmm_nod_0N, type = 3)

# full model minus strain random effect
lmm_nod_2008_dplot <- lmer(sqrt(nodulenum) ~ plantspori + plant_age + plantspori:plant_age + (1|harvester) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2008"))


# model comparison
comp_nod_2008 <- anova(lmm_nod_2008, lmm_nod_2008_dplot)
var_nod_2008 <- VarCorr(lmm_nod_2008)

# test results
chi_nod_2008 <- comp_nod_2008$Chisq
Df_nod_2008 <- comp_nod_2008$Df
sig_nod_2008<- (comp_nod_2008$`Pr(>Chisq)`)/2

# heritability
Var_nod_2008 <- as.data.frame(var_nod_2008)
Var_nod_2008 <- Var_nod_2008 %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_nod_2008 <- Var_nod_2008[1,7]


# combine stats and heritability
comb_nod_2008 <- cbind(chi_nod_2008[2], Df_nod_2008[2], sig_nod_2008[2], VG_nod_2008)
colnames(comb_nod_2008) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")

```

```{r}

#2018
# full model with all random effects
lmm_nod_2018 <- lmer(sqrt(nodulenum) ~ plantspori + plant_age + plantspori:plant_age + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2018"))
Anova(lmm_nod_0N, type = 3)

# full model minus strain random effect
lmm_nod_2018_dplot <- lmer(sqrt(nodulenum) ~ plantspori + plant_age + plantspori:plant_age + (1|harvester) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2018"))


# model comparison
comp_nod_2018 <- anova(lmm_nod_2018, lmm_nod_2018_dplot)
var_nod_2018 <- VarCorr(lmm_nod_2018)

# test results
chi_nod_2018 <- comp_nod_2018$Chisq
Df_nod_2018 <- comp_nod_2018$Df
sig_nod_2018<- (comp_nod_2018$`Pr(>Chisq)`)/2

# heritability
Var_nod_2018 <- as.data.frame(var_nod_2018)
Var_nod_2018 <- Var_nod_2018 %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_nod_2018 <- Var_nod_2018[1,7]


# combine stats and heritability
comb_nod_2018 <- cbind(chi_nod_2018[2], Df_nod_2018[2], sig_nod_2018[2], VG_nod_2018)
colnames(comb_nod_2018) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")

```

# nodulenum heritability by N environment

```{r}

# Control
# full model with all random effects
lmm_nod_0N <- lmer(sqrt(nodulenum) ~ plantspori + nitrogen + plantspori:dataset_year + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(nitrogen=="2008"))
Anova(lmm_nod_0N, type = 3)
summary(lmm_nod_0N)

# model without plot as a random effect
lmm_nod_0N_dplot <- lmer(sqrt(nodulenum) ~ plantspori + dataset_year + plantspori:dataset_year + (1|harvester) + (1|harvestdate), data = metadata %>% filter(nitrogen =="Control"))

# model comparison
comp_nod_0N <- anova(lmm_nod_0N, lmm_nod_0N_dplot)
var_nod_0N <- VarCorr(lmm_nod_0N)

# test results
chi_nod_0N <- comp_nod_0N$Chisq
Df_nod_0N <- comp_nod_0N$Df
sig_nod_0N <- (comp_nod_0N$`Pr(>Chisq)`)/2

# heritability
Var_nod_0N <- as.data.frame(var_nod_0N)
Var_nod_0N <- Var_nod_0N %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_nod_0N <- Var_nod_0N[1,7]


# combine stats and heritability
comb_nod_0N <- cbind(chi_nod_0N[2], Df_nod_0N[2], sig_nod_0N[2], VG_nod_0N)
colnames(comb_nod_0N) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")

```

```{r}
# N-fertilized
# full model with all random effects
lmm_nod_N <- lmer(sqrt(nodulenum) ~ plantspori + dataset_year + plantspori:dataset_year + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(nitrogen=="Nitrogen"))
Anova(lmm_nod_0N, type = 3)
summary(lmm_nod_0N)

# model without plot as a random effect
lmm_nod_N_dplot <- lmer(sqrt(nodulenum) ~ plantspori + dataset_year + plantspori:dataset_year + (1|harvester) + (1|harvestdate), data = metadata %>% filter(nitrogen =="Nitrogen"))

# model comparison
comp_nod_N <- anova(lmm_nod_N, lmm_nod_N_dplot)
var_nod_N <- VarCorr(lmm_nod_N)

# test results
chi_nod_N <- comp_nod_N$Chisq
Df_nod_N <- comp_nod_N$Df
sig_nod_N <- (comp_nod_N$`Pr(>Chisq)`)/2

# heritability
Var_nod_N <- as.data.frame(var_nod_N)
Var_nod_N <- Var_nod_N %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_nod_N <- Var_nod_N[2,7]


# combine stats and heritability
comb_nod_N <- cbind(chi_nod_N[2], Df_nod_N[2], sig_nod_N[2], VG_nod_N)
colnames(comb_nod_N) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")
```



## Rhizobium nodmass models

```{r}
# test if shoot biomass is normally distributed
shapiro.test(metadata$meannodulemass) # p<0.05, transform data
```


# Nodmass heritabiltiy by year

```{r}
# 2008
lmm_nodm_2008 <- lmer(sqrt(meannodulemass) ~ plantspori + plant_age + plantspori:plant_age + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2008"))
Anova(lmm_nodm_2008, type = 3)

# full model minus strain random effect
lmm_nodm_2008_dplot <- lmer(sqrt(meannodulemass) ~ plantspori + plant_age + plantspori:plant_age + (1|harvester) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2008"))


# model comparison
comp_nodm_2008 <- anova(lmm_nodm_2008, lmm_nodm_2008_dplot)
var_nodm_2008 <- VarCorr(lmm_nodm_2008)

# test results
chi_nodm_2008 <- comp_nodm_2008$Chisq
Df_nodm_2008 <- comp_nodm_2008$Df
sig_nodm_2008<- (comp_nodm_2008$`Pr(>Chisq)`)/2

# heritability
Var_nodm_2008 <- as.data.frame(var_nodm_2008)
Var_nodm_2008 <- Var_nodm_2008 %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_nodm_2008 <- Var_nodm_2008[1,7]


# combine stats and heritability
comb_nodm_2008 <- cbind(chi_nodm_2008[2], Df_nodm_2008[2], sig_nodm_2008[2], VG_nodm_2008)
colnames(comb_nodm_2008) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")

```

```{r}

#2018
lmm_nodm_2018 <- lmer(sqrt(meannodulemass) ~ plantspori + plant_age + plantspori:plant_age + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2018"))
Anova(lmm_nodm_2018, type = 3)

# full model minus strain random effect
lmm_nodm_2018_dplot <- lmer(sqrt(meannodulemass) ~ plantspori + plant_age + plantspori:plant_age + (1|harvester) + (1|harvestdate), data = metadata %>%  filter(dataset_year == "2018"))


# model comparison
comp_nodm_2018 <- anova(lmm_nodm_2018, lmm_nodm_2018_dplot)
var_nodm_2018 <- VarCorr(lmm_nodm_2018)

# test results
chi_nodm_2018 <- comp_nodm_2018$Chisq
Df_nodm_2018 <- comp_nodm_2018$Df
sig_nodm_2018<- (comp_nodm_2018$`Pr(>Chisq)`)/2

# heritability
Var_nodm_2018 <- as.data.frame(var_nodm_2018)
Var_nodm_2018 <- Var_nodm_2018 %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_nodm_2018 <- Var_nodm_2018[1,7]


# combine stats and heritability
comb_nodm_2018 <- cbind(chi_nodm_2018[2], Df_nodm_2018[2], sig_nodm_2018[2], VG_nodm_2018)
colnames(comb_nodm_2018) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")

```


# Nodmass heritability by N environment

```{r}
# Control
lmm_nodm_0N <- lmer(sqrt(meannodulemass) ~ plantspori + dataset_year + plantspori:dataset_year + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(nitrogen=="Control"))
Anova(lmm_nodm_0N, type = 3)
summary(lmm_nodm_0N)

# model without plot as a random effect
lmm_nodm_0N_dplot <- lmer(sqrt(meannodulemass) ~ plantspori + dataset_year + plantspori:dataset_year + (1|harvester) + (1|harvestdate), data = metadata %>% filter(nitrogen =="Control"))

# model comparison
comp_nodm_0N <- anova(lmm_nodm_0N, lmm_nodm_0N_dplot)
var_nodm_0N <- VarCorr(lmm_nodm_0N)

# test results
chi_nodm_0N <- comp_nodm_0N$Chisq
Df_nodm_0N <- comp_nodm_0N$Df
sig_nodm_0N <- (comp_nodm_0N$`Pr(>Chisq)`)/2

# heritability
Var_nodm_0N <- as.data.frame(var_nodm_0N)
Var_nodm_0N <- Var_nodm_0N %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_nodm_0N <- Var_nodm_0N[1,7]


# combine stats and heritability
comb_nodm_0N <- cbind(chi_nodm_0N[2], Df_nodm_0N[2], sig_nodm_0N[2], VG_nodm_0N)
colnames(comb_nodm_0N) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")

```

```{r}

# Nitrogen
lmm_nodm_N <- lmer(sqrt(meannodulemass) ~ plantspori + dataset_year + plantspori:dataset_year + (1|harvester) + (1|strain) + (1|harvestdate), data = metadata %>%  filter(nitrogen=="Nitrogen"))
Anova(lmm_nodm_N, type = 3)
summary(lmm_nodm_N)

# model without plot as a random effect
lmm_nodm_N_dplot <- lmer(sqrt(meannodulemass) ~ plantspori + dataset_year + plantspori:dataset_year + (1|harvester) + (1|harvestdate), data = metadata %>% filter(nitrogen =="Nitrogen"))

# model comparison
comp_nodm_N <- anova(lmm_nodm_N, lmm_nodm_N_dplot)
var_nodm_N <- VarCorr(lmm_nodm_N)

# test results
chi_nodm_N <- comp_nodm_N$Chisq
Df_nodm_N <- comp_nodm_N$Df
sig_nodm_N <- (comp_nodm_N$`Pr(>Chisq)`)/2

# heritability
Var_nodm_N <- as.data.frame(var_nodm_N)
Var_nodm_N <- Var_nodm_N %>%
  mutate(VP = sum(vcov),
         prop = vcov/VP)

VG_nodm_N <- Var_nodm_N[1,7]


# combine stats and heritability
comb_nodm_N <- cbind(chi_nodm_N[2], Df_nodm_N[2], sig_nodm_N[2], VG_nodm_N)
colnames(comb_nodm_N) <- c("Chisq", "Df", "p-value", "broad-sense-heritability")
```




__